R and Python clients for GBIF species occurrence data

### 2017-08-10 manuscript.Rmd

`rgbif` and `pygbif` are designed following the [GBIF Application Programming Interface][gbifdocs],
or API. The GBIF API has four major components: registry, taxonomic names,
occurrences, and maps. We also include functions to interface with the OAI-PMH
GBIF service; only dataset information is available vis this service, however. We
ignore maps in both libraries as it is concerned with generating maps primarily for web
applications. Both libraries have a suite of functions dealing with each of registry, taxonomic
names, and occurrences - we'll go through each in turn describing design and example usage.

@@

`pygbif` only has `name_backbone()` and `name_suggest()` at this time.

@@

@@

The `rgbif` and `pygbif` libraries provide programmatic interfaces to GBIF's
application programming interface (API) - a powerful tool for working with
species occurrence data, and facilitating reproducible research. In fact,
the `rgbif` package has already been used in 22 scholarly publications
(as of 2015-11-14).

>>>>

`rgbif` and `pygbif` are designed following the [GBIF Application Programming Interface][gbifdocs],
or API. The GBIF API has four major components: registry, taxonomic names,
occurrences, and maps. We also include functions to interface with the OAI-PMH
GBIF service; only dataset information is available vis this service, however. We now include
an interface to the GBIF map API, but we for now ignore maps in `pygbif`. Both libraries have a
suite of functions dealing with each of registry, taxonomic, names, and occurrences - we'll go
through each in turn describing design of the user interface and example usage.

@@

`pygbif` has all the same method except `rgbif::parsenames()` is `name_parser()`
in `pygbif`.

@@

## GBIF Mapping API

<!-- FILL ME IN -->

@@

The `rgbif` and `pygbif` libraries provide programmatic interfaces to GBIF's
application programming interface (API) - a powerful tool for working with
species occurrence data, and facilitating reproducible research. In fact,
the `rgbif` package has already been used in 20 scholarly publications
(as of 2016-08-10) [@Amano_2016, @Bartomeus_2013, @Barve_2014, @Bone_2015, @Collins_2015, @Drozd_2013, @Kong_2015, @Richardson_2015, @Turner_2015, @Verheijen_2015, @Zizka_2015, @Butterfield_2016, @Dellinger_2015, @Feitosa_2015, @Malhado_2015, @Werner_2015, @Robertson_2016, @Davison_2015, @Janssens_2016].

### 2017-08-10 use-cases/oak-occurrences.Rmd

From [http://esp.cr.usgs.gov/data/little/](http://esp.cr.usgs.gov/data/little/). And save shapefile in same directory

```{r}
url <- 'http://esp.cr.usgs.gov/data/little/querloba.zip'
tmp <- tempdir()
download.file(url, destfile = "~/querloba.zip")
unzip("~/querloba.zip", exdir = tmp)
ql <- readShapePoly(file.path(tmp, "querloba.shp"))
```

@@

```{r}
alt.USA <- getData('alt', country = 'USA')
```

@@

```{r}
alt.USA <- alt.USA[[1]]
slope.USA <- terrain(alt.USA, opt = 'slope')
aspect.USA <- terrain(alt.USA, opt = 'aspect')
hill.USA <- hillShade(slope.USA, aspect.USA, angle = 45, direction = 315)
```

@@

```{r fig.keep='last'}
plot(hill.USA, col = grey(0:100/100), legend = FALSE, xlim = c(-125, -116), ylim = c(32, 42), main = 'Distribution of Quercus lobata', xlab = "Long", ylab = 'Lat')
# add shape from Atlas of US Trees
plot(ql, add = TRUE, col = alpha("white", 0.6), border = FALSE)
# add Gbif presence points
points(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude, cex = .7, pch = 19, col = alpha("darkgreen", 0.8))
legend(x = -121, y = 40.5, "GBIF Data", pch = 19, col = 'darkgreen', bty = 'n', pt.cex = 1, cex = .8)
legend(x = -121, y = 41.5, "Atlas of United States Trees \n (Little, E. 1971)", pt.cex = 1.5, cex = .8, pch = 19, col = 'white', bty = 'n')
```

>>>>

From <http://esp.cr.usgs.gov/data/little>. And save shapefile in same directory

```{r cache = TRUE}
url <- 'http://esp.cr.usgs.gov/data/little/querloba.zip'
tmp <- tempdir()
download.file(url, destfile = "~/querloba.zip")
unzip("~/querloba.zip", exdir = tmp)
ql <- readShapePoly(file.path(tmp, "querloba.shp"))
```

@@

```{r cache = TRUE}
alt.USA <- getData('alt', country = 'USA')
```

@@

```{r cache = TRUE}
alt.USA <- alt.USA[[1]]
slope.USA <- terrain(alt.USA, opt = 'slope')
aspect.USA <- terrain(alt.USA, opt = 'aspect')
hill.USA <- hillShade(slope.USA, aspect.USA, angle = 45, direction = 315)
```

@@

```{r fig.keep='last'}
plot(hill.USA, col = grey(0:100/100), legend = FALSE,
     xlim = c(-125, -116), ylim = c(32, 42),
     main = 'Distribution of Quercus lobata',
     xlab = "Long", ylab = 'Lat')
# add shape from Atlas of US Trees
plot(ql, add = TRUE, col = alpha("white", 0.6),
     border = FALSE)
# add Gbif presence points
points(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude,
       cex = .7, pch = 19, col = alpha("darkgreen", 0.8))
legend(x = -121, y = 40.5, "GBIF Data", pch = 19,
       col = 'darkgreen', bty = 'n', pt.cex = 1, cex = .8)
legend(x = -121, y = 41.5,
       legend = "Atlas of United States Trees \n (Little, E. 1971)",
       pt.cex = 1.5, cex = .8, pch = 19, col = 'white', bty = 'n')
```
